Using Occupancy Theory and Hierarchical Models to Estimate Otter Population in Southeastern Minnesota

### 2010-02-25 Paper.tex
Added tables and figure.

    Simulations of \(S\) show that the distribution of \(S\) is roughly normal.
    So using our calculations for the mean and the standard deviation, we
    computed a Z-score for each flight, where we use a plug-in estimate for
    \(p\) to be the proportion of \(1\)'s per flight. We said that Z-scores with
    an absolute value greater than \(2\) suggest that the assumption of
    independence between sites is violated for that flight. We found that, in
    general, alternating sites are more "independent" than all sites as measured
    by this statistic. However, alternating sites do not guarantee independence.
    For instance, six of eleven flights along the Zumbro River during 2003 still
    have significant Z-scores with alternating sites. There were no noticeable
    trends for this statistic with respect to observer, days since snow, snow
    event, or plot size.

    % FIGURES OR TABLES??? WHICH ONES? from status report compilation

@@

    \subsection{Our Hierarchical Model}
    For our model, we specified uniform \((0, 1)\) prior distributions for
    parameters \(\psi\) and \(\theta\), a common, natural choice given a lack of
    formal information about the prior or expert consensus \cite{MacKenzie2006}.
    We implemented models with normal and beta prior distributions on \(p\) and
    \(E\) and limited the range of \(p\) to the interval \((0.6, 1)\) and the
    range of \(E\) to the interval \((0, 0.3)\) in both models. We set the
    burn-in time for the MCMC simulation to 10,000 iterations. Each of these
    parameters are tracked and therefore are estimated.

>>>>

    Simulations of \(S\) show that the distribution of \(S\) is roughly normal.
    So using our calculations for the mean and the standard deviation, we
    computed a Z-score for each flight, where we use a plug-in estimate for
    \(p\) to be the proportion of \(1\)'s per flight. We said that Z-scores with
    an absolute value greater than \(2\) suggest that the assumption of
    independence between sites is violated for that flight. We found that, in
    general, alternating sites are more "independent" than all sites as measured
    by this statistic. However, alternating sites do not guarantee independence.
    For instance, six of eleven flights along the Zumbro River during 2003 still
    have significant Z-scores with alternating sites. There were no noticeable
    trends for this statistic with respect to observer, days since snow, snow
    event, or site size.

    % Mississippi River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|l|}
        \hline
        \multicolumn{6}{|l|}{\textbf{Mississippi River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} &
        \multirow{2}{*}{\(E[|z_{all}|-|z_{alt}|]\)} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots & \\
        \hline
        \multirow{3}{*}{2003} & 402m & \multirow{3}{*}{20} & 7 & 0 & 0.915 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 5 & 0 & 0.651 \\
        \cline{2-2} \cline{4-6}
        & 1608m & & 1 & 0 & 0.383 \\
        \hline
        \multirow{3}{*}{2004} & 402m & \multirow{3}{*}{6} & 5 & 3 & 0.625 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 3 & 0 & 1.428 \\
        \cline{2-2} \cline{4-6}
        & 1608m & & 2 & 0 & 0.053 \\
        \hline
    \end{tabular}
    \end{center}

    % Whitewater River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|l|}
        \hline
        \multicolumn{6}{|l|}{\textbf{Whitewater River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} &
        \multirow{2}{*}{\(E[|z_{all}|-|z_{alt}|]\)} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots & \\
        \hline
        \multirow{2}{*}{2003} & 402m & \multirow{2}{*}{15} & 6 & 2 & 0.722 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 4 & 1 & 0.634 \\
        \hline
        \multirow{2}{*}{2004} & 402m & \multirow{2}{*}{6} & 3 & 1 & 1.189 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 2 & 2 & 0.612 \\
        \hline
    \end{tabular}
    \end{center}

    % Zumbro River!
    \begin{center}
    \begin{tabular}{|l|l|l|l|l|l|}
        \hline
        \multicolumn{6}{|l|}{\textbf{Zumbro River}} \\
        \hline
        \multirow{2}{*}{Year} & \multirow{2}{*}{Size} & \multirow{2}{*}{Flights}
        & \multicolumn{2}{|l|}{\# Sig. Z-Scores} &
        \multirow{2}{*}{\(E[|z_{all}|-|z_{alt}|]\)} \\
        \cline{4-5}
        & & & All Plots & Alt. Plots & \\
        \hline
        \multirow{2}{*}{2003} & 402m & \multirow{2}{*}{11} & 10 & 6 & 1.883 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 8 & 5 & 1.287 \\
        \hline
        \multirow{2}{*}{2004} & 402m & \multirow{2}{*}{4} & 3 & 3 & 2.425 \\
        \cline{2-2} \cline{4-6}
        & 804m & & 3 & 1 & 2.343 \\
        \hline
    \end{tabular}
    \end{center}

@@

    \subsection{Our Hierarchical Model}
    For our model, we specified uniform \((0, 1)\) prior distributions for
    parameters \(\psi\) and \(\theta\), a common, natural choice given a lack of
    formal information about the prior or expert consensus \cite{MacKenzie2006}.
    We implemented models with normal and beta prior distributions on \(p\) and
    \(E\) and limited the range of \(p\) to the interval \((0.6, 1)\) and the
    range of \(E\) to the interval \((0, 0.4)\) in both models. We set the
    burn-in time for the MCMC simulation to 10,000 iterations. Each of these
    parameters are tracked and therefore are estimated. The covariates of this
    model are: observer, date, and the number of days since the last snow event.
